### 2010-07-23 trunk/textos/TI2/texto/03-bezerra2009lbs.tex

In \cite{devleeschauwer2005dma}, for example, a few algorithms are proposed. Their idea is to transfer regions managed by overloaded servers to  lightly loaded ones, in a way such that the maximum overload among the servers is reduced. Some very interesting principles are used, as the concepts of microcells and macrocells (which will be described in the next section). However, their model has some differences to the one presented in this work. For instance, they consider the number of players as the load to distribute, heterogeneous systems are not supported, and the algorithm they use to refine an existing partitioning is simulated annealing. This one may be a little too general, considering that there are algorithms specific to partitioning, such as \cite{kernighan1970ehp} or \cite{fiduccia1982lth}, which may perform better. Finally, some of the algorithms proposed in \cite{devleeschauwer2005dma} do not consider the communication overhead between servers.

@@

Chen et al. propose, in \cite{chen2005lad}, a load shedding technique, in which an overloaded server attempts do shed its load to its neighbors. After finding a lightly loaded neighbor server -- if there is one -- it transfers some of its boundary microcells to that neighbor. To form the group of microcells to transfer, a microcell from the border is chosen, and others are added in a breadth-first search (BFS) order. A very similar work, which also uses BFS, is described in \cite{duong2003dls}. However, these models present a few points which could be improved. For example, a server is considered overloaded when a certain number $n$ of players is reached. As we said before, this is not the best measure, since the load on a server may vary completely, depending on how these $n$ players are distributed across the region managed by the server. Also, the use of BFS to merge two partitions or split one of them may not be ideal. Some other algorithm for graph partitioning applied to distributed systems \cite{kernighan1970ehp,fiduccia1982lth,karypis1999fah,hendrickson1995isg} could be used, for most of these algorithms were designed exactly to balance load and minimize dependence between nodes.

@@

In the previous section it was presented some existing works regarding load balancing in MMOGs when using multiple servers to provide the support network. The load balancing scheme proposed in this work is based on some of the principles in the literature. One of them is the division of the virtual environment in microcells, for later grouping in macrocells. This is a relatively simple way of addressing the issue of the avatars' movement dynamics through the game world, by transferring the microcells dynamically according to need.

It will also be used the idea of balancing based only on local information -- each server, when needing to reduce its workload, selects only a few other servers to join a local load rebalancing. Thus, we can greatly reduce the complexity of balancing because it will not be necessary that all servers in the game exchange messages among themselves every time that any one of them is overloaded.

To define our load balancing approach, we cannot ignore that, usually, the traffic generated by the players is not simply linear, but square for each cluster of players. Such misunderstanding can generate considerable differences between the actual load of each server and the load estimated by the balancing algorithm. Another point to consider is the overhead, both in the delay of sending messages, as in the use of bandwidth of the servers, when players connected to different servers are interacting with each other. Some works assume that the servers are all in the same high-speed and low latency local-area network, and therefore overhead is negligible. However, when considering a geographically distributed server system, this assumption cannot be made. This overhead must be taken into account, no matter which load balancing algorithm is being used.

Another important point is that the main criterion we use when balancing the load of MMOG servers is the bandwidth, and not processing power. Several games, such as Age of Empires \cite{ageofempires}, include simulations of virtual environments with hundreds or thousands of entities, which are performed smoothly on current personal computers. However, if this game was multiplayer and each of these entities was controlled by a different player on the Internet, it would most likely generate a traffic amount which would hardly be supported by a domestic connection \cite{feng2007wnn}.

@@

There are also other issues which must not be overlooked. One of them is that the server system is probably not homogenous -- considering that it is geographically distributed with, likely, different connections to the Internet. Therefore, one cannot assume that the servers have the same amount of cpu power or network bandwidth. % Another problem, specific to the algorithm proposed by \cite{lee2003sdl}, is that the criterion to allocate to each server a load equivalent to 90\% of its capacity in order to avoid constant rebalancing is weak. There are no guarantees that the server system has 11.11\% more total capacity than the necessary. On the contrary, it should adapt to situations of widespread overload. Moreover, the criterion for stopping is weak because the algorithm does not stop when the whole system is loaded above 90\% of its capacity, or if the overloaded server has no neighbors capable of receiving more load.
In the next section, we make some definitions that will be necessary to specify our load balancing scheme.

@@

We also use the idea of mapping the virtual environment on a graph. The graph will then be partitioned to distribute the workload of the game between the different servers. It is necessary first to define some terms which will be used on the proposed algorithms.

@@

Figure \ref{fig:mapping}(a) illustrates how the mapping is done with square cells, while figure \ref{fig:mapping}(b) shows how it would be with hexagonal cells. The objective of the balancing scheme proposed here is to assign to each server a weight proportional to its capacity, reducing as much as possible the edge-cut of the graph that represents the virtual environment and, consequently, reduce the overhead inherent to the distribution of the game on multiple servers. Although this is an NP-complete problem \cite{feder1999cgp}, efficient heuristics will be used to reduce this overhead. In the following sections the algorithms proposed in this work will be presented.

@@

Like we said, it is passed as input a list of the regions whose load will be rebalanced. This makes possible the use of this algorithm both in local and global scope, just by choosing between passing some regions or all regions of the environment. The distribution is based on information from that set of regions, whose total weight and total power are calculated in lines \ref{alg:ggp:begingetlocalinfo} to \ref{alg:ggp:endgetlocalinfo}. To be redistributed later, all cells associated with these regions are released (line \ref{alg:ggp:freeallcells}).

@@

A possible undesirable effect of the \ggp\ algorithm is that by releasing all the cells and redistributing them, it might happen that one or more regions completely change their place, causing several players to disconnect from their servers and reconnect to a new one. To try to reduce the likelihood of such event, we propose a variation of the algorithm, called \textbf{\ggpk} (\ggpkmeaning). This new algorithm (shown in Algorithm \ref{alg:ggpk}) is similar to the original version, except that each region maintains its heaviest cell (lines \ref{alg:ggpk:saveheaviestcell} and \ref{alg:ggpk:getbackheaviestcell}), from which a region similar to the previous one can be formed, so that several players will not need to migrate to other server. However, to keep one of the cells of each region, it might be preventing a better balancing to occur. Also, the fixation of that cell might cause fragmented regions, increasing the overhead.

@@

The \textbf{\bfa} (\bfameaning) is proposed here as an alternative to \ggp\ and its variants. The objective of the algorithm is to check what is the weight excess in each region and transfer it to free regions whose capacity is the closest to that value. This is done by transferring cells whose weight is the closest to free capacity of the receiving region, observed two restrictions: first, the total weight transferred can not be larger than the free capacity of the destination region and, second, we should not transfer a load greater than the necessary to eliminate the overload. The second restriction is justified because a weight transfer larger than necessary would probably result in a larger amount of migrating players. The exception to this rule is when a cell is heavier than the other, not for having more avatars, but because they are closer, with quadratic traffic growth between them. Algorithm \ref{alg:bfa} describes in detail the operation of \bfa.

@@

To attenuate this problem, it is proposed here to use the algorithm of Kernighan and Lin \cite{kernighan1970ehp}. This algorithm receives as input the graph that represents the virtual environment. Given a pair of regions (partitions), \kl\ searches for pairs of cells (vertices) that, when exchanged between their regions, the overhead (edge-cut) is reduced.

>>>>

In \cite{devleeschauwer2005dma}, for example, a few algorithms are proposed. Their idea is to transfer regions managed by overloaded servers to  lightly loaded ones, in a way such that the maximum overload among the servers is reduced. Some very interesting principles are used, as the concepts of microcells and macrocells (which will be described in the next section). However, their model has some differences to the one presented here. For instance, they consider the number of players as the load to distribute, heterogeneous systems are not supported, and the algorithm they use to refine an existing partitioning is simulated annealing. This one may be a little too general, considering that there are algorithms specific to partitioning, such as \cite{kernighan1970ehp} or \cite{fiduccia1982lth}, which may perform better. Finally, some of the algorithms proposed in \cite{devleeschauwer2005dma} do not consider the communication overhead between servers.

@@

Chen et al. propose, in \cite{chen2005lad}, a load shedding technique, in which an overloaded server attempts do shed its load to its neighbors. After finding a lightly loaded neighbor server -- if there is one -- it transfers some of its boundary microcells to that neighbor. To form the group of microcells to transfer, a microcell from the border is chosen, and others are added in a breadth-first search (BFS) order. A very similar work, which also uses BFS, is described in \cite{duong2003dls}. However, these models present a few points which could be improved. For example, a server is considered overloaded when a certain number $n$ of players is reached. As mentioned before, this is not the best measure, since the load on a server may vary completely, depending on how these $n$ players are distributed across the region managed by the server. Also, the use of BFS to merge two partitions or split one of them may not be ideal. Some other algorithm for graph partitioning applied to distributed systems \cite{kernighan1970ehp,fiduccia1982lth,karypis1999fah,hendrickson1995isg} could be used, for most of these algorithms were designed exactly to balance load and minimize dependence between nodes.

@@

In the previous section it was presented some existing works regarding load balancing in MMOGs when using multiple servers to provide the support network. The load balancing scheme presented here is based on some of the principles in the literature. One of them is the division of the virtual environment in microcells, for later grouping in macrocells. This is a relatively simple way of addressing the issue of the avatars' movement dynamics through the game world, by transferring the microcells dynamically according to need.

It will also be used the idea of balancing based only on local information -- each server, when needing to reduce its workload, selects only a few other servers to join a local load rebalancing. Thus, the complexity of balancing can be greatly reduced because it will not be necessary for all servers in the game to exchange messages among themselves every time that any one of them is overloaded.

To define their load balancing approach, the authors consider that, usually, the traffic generated by the players is not simply linear, but square for each cluster of players. Such misunderstanding can generate considerable differences between the actual load of each server and the load estimated by the balancing algorithm. Another point to consider is the overhead, both in the delay of sending messages, as in the use of bandwidth of the servers, when players connected to different servers are interacting with each other. Some works assume that the servers are all in the same high-speed and low latency local-area network, and therefore overhead is negligible. However, when considering a geographically distributed server system, this assumption cannot be made. This overhead must be taken into account, no matter which load balancing algorithm is being used.

Another important point is that the main criterion the authors use when balancing the load of MMOG servers is the bandwidth, and not processing power. Several games, such as Age of Empires \cite{ageofempires}, include simulations of virtual environments with hundreds or thousands of entities, which are performed smoothly on current personal computers. However, if this game was multiplayer and each of these entities was controlled by a different player on the Internet, it would most likely generate a traffic amount which would hardly be supported by a domestic connection \cite{feng2007wnn}.

@@

There are also other issues which must not be overlooked. One of them is that the server system is probably not homogenous -- considering that it is geographically distributed with, likely, different connections to the Internet. Therefore, one cannot assume that the servers have the same amount of cpu power or network bandwidth. % Another problem, specific to the algorithm proposed by \cite{lee2003sdl}, is that the criterion to allocate to each server a load equivalent to 90\% of its capacity in order to avoid constant rebalancing is weak. There are no guarantees that the server system has 11.11\% more total capacity than the necessary. On the contrary, it should adapt to situations of widespread overload. Moreover, the criterion for stopping is weak because the algorithm does not stop when the whole system is loaded above 90\% of its capacity, or if the overloaded server has no neighbors capable of receiving more load.
In the next section, it is made some definitions that will be necessary to specify the load balancing scheme proposed by the \cite{bezerra2009lbs}.

@@

The authors also use the idea of mapping the virtual environment on a graph. The graph will then be partitioned to distribute the workload of the game between the different servers. It is necessary first to define some terms which will be used on the proposed algorithms.

@@

Figure \ref{fig:mapping}(a) illustrates how the mapping is done with square cells, while figure \ref{fig:mapping}(b) shows how it would be with hexagonal cells. The objective of the balancing scheme presented here is to assign to each server a weight proportional to its capacity, reducing as much as possible the edge-cut of the graph that represents the virtual environment and, consequently, reduce the overhead inherent to the distribution of the game on multiple servers. Although this is an NP-complete problem \cite{feder1999cgp}, efficient heuristics will be used to reduce this overhead. In the following sections the algorithms proposed by the authors will be presented.

@@

Like said before, it is passed as input a list of the regions whose load will be rebalanced. This makes possible the use of this algorithm both in local and global scope, just by choosing between passing some regions or all regions of the environment. The distribution is based on information from that set of regions, whose total weight and total power are calculated in lines \ref{alg:ggp:begingetlocalinfo} to \ref{alg:ggp:endgetlocalinfo}. To be redistributed later, all cells associated with these regions are released (line \ref{alg:ggp:freeallcells}).

@@

A possible undesirable effect of the \ggp\ algorithm is that by releasing all the cells and redistributing them, it might happen that one or more regions completely change their place, causing several players to disconnect from their servers and reconnect to a new one. Trying to reduce the likelihood of such event, the authors propose a variation of the algorithm, called \textbf{\ggpk} (\ggpkmeaning). This new algorithm (shown in Algorithm \ref{alg:ggpk}) is similar to the original version, except that each region maintains its heaviest cell (lines \ref{alg:ggpk:saveheaviestcell} and \ref{alg:ggpk:getbackheaviestcell}), from which a region similar to the previous one can be formed, so that several players will not need to migrate to other server. However, to keep one of the cells of each region, it might be preventing a better balancing to occur. Also, the fixation of that cell might cause fragmented regions, increasing the overhead.

@@

The \textbf{\bfa} (\bfameaning) was proposed by the authors as an alternative to \ggp\ and its variants. The objective of the algorithm is to check what is the weight excess in each region and transfer it to free regions whose capacity is the closest to that value. This is done by transferring cells whose weight is the closest to free capacity of the receiving region, observed two restrictions: first, the total weight transferred can not be larger than the free capacity of the destination region and, second, it should not be transferred a load greater than the necessary to eliminate the overload. The second restriction is justified because a weight transfer larger than necessary would probably result in a larger amount of migrating players. The exception to this rule is when a cell is heavier than the other, not for having more avatars, but because they are closer, with quadratic traffic growth between them. Algorithm \ref{alg:bfa} describes in detail the operation of \bfa.

@@

To attenuate this problem, it was also proposed by \cite{bezerra2009lbs} to use the algorithm of Kernighan and Lin \cite{kernighan1970ehp}. This algorithm receives as input the graph that represents the virtual environment. Given a pair of regions (partitions), \kl\ searches for pairs of cells (vertices) that, when exchanged between their regions, the overhead (edge-cut) is reduced.
